Lyman-$\alpha$ photons through rotating outflows

### 2018-09-10 paper/paper.tex
edit

\begin{abstract}
Outflows and rotation are two ubiquitous kinematic features in the gas
kinematics of galaxies. 
Here we introduce a semi-analytic model to quantify how rotating
outflows impact the morphology of the Lyman-$\alpha$ emission line.   
The model is contrasted against Monte Carlo radiative transfer
simulations of outflowing gas with additional solid body rotation.
We explore a range of neutral Hydrogen optical depth of
$10^5\leq\tauh\leq10^7$, rotational velocity $0\leq \vrot/\kms \leq
100$ and outflow velocity $0\leq \vout/\kms\leq 50$.  
We find three consequences of rotation.
First, it introduces a dependency with viewing angle; second it
produces a line broadening and third it increases the flux at the
line's center.
For all viewing angles, the semi-analytic model reproduces the
radiative transfer results for the line width and flux change at the
line's center within a $7\%$ and $50\%$ precision for an optical depth of
$\tauh=10^5$, respectively, and within $2\%$ and $1\%$ for an optical
depth of $\tauh=10^7$.
Using this toy model we also show that the peaks of integrated spectra
taken from opposite sides of an edge-on rotating gas distribution
should have a separation of $\frac{1}{2}\vrot$. 
The semi-analytic model presented here is a convenient tool to
introduce rotational kinematics as a post-processing step of idealized
Monte Carlo simulations, providing a framework to interpret \lya
spectra in systems where rotation is expected, or directly measured
through kinematic maps.  
\end{abstract}

@@

In this paper we explore, for the first time in the literature,
the results of a model for the emergent \lya line from rotating
outflows. 
We use a semi-analytic model, first presented in \citep{Garavito14}, 
to capture the main effects of rotation and confront it against
results from Monte-Carlo radiative transfer simulations. 
The semi-analytic model only takes into account the Doppler shift computed
as product of quantities at the surface of last scattering, namely
$\vec{v}_{\rm rot}\cdot \hatk$, where $\vec{v}_{\rm rot}$ is the
velocity due to rotation and \hatk is the direction of the photon's
propagation.

To address the first question we posed in the introduction (\emph{what is then
the expected imprint of rotation on a resonant emission line such as
the \lya line?}) we can say that the main effects of rotation on the
\lya line morphology are:

\begin{itemize}
  \item Inducing a dependency on the viewing angle.
  \item Broadening the line.
  \item Increasing the intensity at the line's center.
\end{itemize}
%
All these effects can be quantitatively explained by the proposed
semi-analytic model.

Addressing the second question we posed in the introduction (\emph{to what
extent is it possible to constrain rotational kinematics from the \lya
emission line?})  we can say that  the most straightforward approach
takes into account how the \lya spectra from two different sides of a
galaxy can detect approaching/receding gas motions
\citep{Prescott14,2018MNRAS.473.3907A}. 
In that case, we also show that one has to be careful and take into
account that the  peaks of these two spectra is smaller than a naively
expected value of $2\times\vrot$ due to the different weights at the
surface of the emitting regions.   
Here the semi-analytic model also provides the correct quantitative
insight.

To summarize, our work confirms that the semi-analytic model based on
the Doppler Shift offers an easy-to-implement approximation to explore
such influence into already existing radiative transfer simulation
simulations and provide stronger footing to the interpretation of
present and future \lya kinematics maps
\citep[e.g][]{2018MNRAS.473.3907A}.

>>>>

\begin{abstract}
Outflows and rotation are two ubiquitous kinematic features in the gas
kinematics of galaxies. 
Here we introduce a semi-analytic model to quantify how rotating
outflows impact the morphology of the Lyman-$\alpha$ emission line.   
The model is contrasted against Monte Carlo radiative transfer
simulations of outflowing gas with additional solid body rotation.
We explore a range of neutral Hydrogen optical depth of
$10^5\leq\tauh\leq10^7$, rotational velocity $0\leq \vrot/\kms \leq
100$ and outflow velocity $0\leq \vout/\kms\leq 50$.  
We find three consequences of rotation.
First, it introduces a dependency with viewing angle; second it
broadens the line and third it increases the flux at the
line's center.
For all viewing angles, the semi-analytic model reproduces the
radiative transfer results for the line width and flux change at the
line's center within a $7\%$ and $50\%$ precision for an optical depth of
$\tauh=10^5$, respectively, and within $2\%$ and $1\%$ for an optical
depth of $\tauh=10^7$.
Using this model we also show that the peaks of integrated spectra
taken from opposite sides of an edge-on rotating gas distribution
should have a separation of $\frac{1}{2}\vrot$. 
The semi-analytic model presented here is a convenient tool to
introduce rotational kinematics as a post-processing step of idealized
Monte Carlo simulations; it provides a framework to interpret \lya
spectra in systems where rotation is expected or directly measured
through kinematic maps.  
\end{abstract}

@@

In this paper we explore, for the first time in the literature,
the results of a model for the emergent \lya line from rotating
outflows. 
We the a semi-analytic model first presented by \cite{Garavito14}
to capture the main effects of rotation and confront it against
results from Monte-Carlo radiative transfer simulations. 
The semi-analytic model only takes into account the Doppler shift computed
as product of quantities at the surface of last scattering, namely
$\vec{v}_{\rm rot}\cdot \hatk$, where $\vec{v}_{\rm rot}$ is the
velocity due to rotation and \hatk is the direction of the photon's
propagation.

To address the first question we posed in the introduction (\emph{What
  is the expected imprint of rotation on a resonant emission line such
  as the \lya line?}) we find from the full RT simulations that the
effects of rotation on the \lya line morphology are:

\begin{itemize}
  \item Inducing a dependency on the viewing angle.
  \item Broadening the line.
  \item Increasing the intensity at the line's center.
\end{itemize}
%
All these effects can be qualitatively explained by the proposed
semi-analytic model. 
Quantitatively speaking the semi-analytic model provides a
satisfactory answer for a neutral Hydrogen optical depth equal or
larger than $10^6$.

Addressing the second question we posed in the introduction (\emph{To what
extent is it possible to constrain rotational kinematics from the \lya
emission line?})  we show that the most straightforward approach
is measuring \lya spectra from two different regions of a
galaxy to detect approaching/receeding gas motions
\citep{Prescott14,2018MNRAS.473.3907A}. 
In that case, we also show that the distances between these two peaks 
is four times shorter than the naively expected value of
$2\times\vrot$.  
This difference is produced by the different geometric weights at the
emitting surface.    
Here the semi-analytic model also provides the correct quantitative
insight.

To summarize, our works shows that the Doppler Shift offers an
easy-to-implement approximation to explore such influence into already
existing radiative transfer simulation simulations that include
outflows, providing a versatile tool to interpret current and future \lya
kinematics maps  \citep[e.g][]{2018MNRAS.473.3907A}.
